Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-22 demystify.tex
updates

\subsection{Self-written Letters are Bad}\label{sec:self-letters}
 Many international students write letters themselves, typically due to the requests of their profs. or letter writers, and have them signed by their profs. Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?  Instead of the ref. writer talking about you, in this case it is you who write about yourself (and they just sign the letter). Experienced reviewers can easily spot  (because the student often has little experience in writing LoRs) and  \red{red flag} such letters.

Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this). This might be a common practice at specific universities and the students do not have a choice as they need the letter.  However, think about this: if a prof. does this often, then they either do not know/care enough about you or don't know how to write a good letter.  In any case, such letters are not useful and might even hurt your application.  So if you are in this situation, you should find someone else to write for you.

\subsection{Letters from Admin People}\label{sec:admin-letters}

Many international students try to get letters from high-ranking admin people in their universities such as department chair/head, dean, or director. The students often believe that these letters are more valuable because such people are important in their universities.  However, these people, while well-known and respected in your university, might not be active in research (e.g., they focus on administration) and thus are not recognized by adcom members. Such a letter has little value if the writer does not know you well enough (e.g., did not work with you) to talk about your research ability.

\begin{commentbox}[Vu:]
  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.  So if you are in this situation, you should find someone else to write for you.
    So if you are in this situation, find someone else to write for you.
\end{commentbox}

>>>>

\subsection{Self-written Letters are Bad}\label{sec:self-letters}
 Many international students write letters themselves, typically due to the requests of their profs. or letter writers. Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?  Instead of the ref. writer talking about you, in this case it is you who write about yourself (and they just sign the letter). Experienced reviewers can easily spot  (because the student often has little experience in writing LoRs) and  \red{red flag} such letters.

Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about). This might be a common practice at specific universities and the students do not have a choice as they need the letter.  However, think about this: if a prof. does this often, then they either don't know how to write a LOR (more common than you would think) or simply do not know or care enough about you.  In any case, such LoRs are not useful and might even hurt your application.  So if you are in this situation, you should find someone else to write for you.

\subsection{LoRs from Admin People}\label{sec:admin-letters}

Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director. The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
However, as emphasized, a generic LoR has little value because the writer does not know you well and can talk in depth about your research ability.

Moreover, while being well-known and respected in your local university, these writers might not be very active in research (e.g., they haven't published in recent years). Thus they might not be well-known and recognized by adcom members.

\begin{commentbox}[Vu:]
  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.  So if you are in this situation, you should find someone else to write for you as mentioned in this section.
\end{commentbox}
